---
title: "MATH1115 Computer Practical Exam (SAMPLE TEMPLATE)"
subtitle: "Semester 1 2019"
date: 
output:
  html_document:
    number_sections: yes
    self_contained: yes
    theme: flatly 
    toc: true  # Table of contents
    toc_depth: 3
    toc_float: true
---

<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

*Note: below is an empty template showing you the basic format of your exam. You will be given a template like this, populated with a url for the data, and research questions.*

## MY SID IS: XXXXX {-}

<br>

## INITIAL DATA ANALYSIS

Load packages and data
```{r, warning=F, message=F}
library(tidyverse)
library(ggthemes)
library(ggplot2)
#file = read.csv(file = "url given")
```

**Question1**: 

```{r}
# Put your code here
```

<>

**Question2**: 

```{r}
# Add code here
```

<>

**Question3**: 

```{r}
# Add code here
```

<>

<br>

## SIMPLE LINEAR REGRESSION

**Research Question**: 

```{r}
# Add code here
```

<>

<br>

## CHI-SQUARED TEST

**Research Question**: 

```{r}
# Put your code here
```

<>

<br>

## BINOMIAL MODELLING

Context:

```{r}
# Add Code Here
```

<>

<br>

## DATA WRANGLING

Annotate the code below, so that it explains what each line is doing.

```{r, warning = F, message = F, eval = F}
# code given
```


